\specialchapt{ABSTRACT}

This thesis explores using a software defined radio (SDR) to create a SDR-based radiometer that is capable of performing the same operation as a traditional radiometer and offers additional capabilities not found in traditional radiometers.  Traditional radiometer requires careful design to ensure correct operation, by digitizing the RF signal as soon as possible and processing this signal in software, the hardware design of the radiometer can be simplified.

Digital radiometers have been explored before, but often use customized components.  Software defined radio technology has become more widespread, and affordable Commercial Off The Shelf (COTS) SDRs are now available with high performance.  This thesis leverages  a COTS SDR technology to implement and evaluate a SDR-based radiometer.  This will lower the cost of the radiometer and help make radiometers more accessible to a wider audience.  The mapping of the functionality of a traditional radiometer to our proposed SDR-based radiometer is examined.  Then an experimental evaluation of the performance between a traditional and SDR-based radiometer is conducted.  Additionally this thesis explores how the implemented SDR-based radiometer can help mitigate radio frequency interference.

%Discussed in this thesis is how parts of a traditional radiometer is emulated in software.  This is then verified through experimentation that compares data collected from a SDR-based radiometer and a traditional radiometer.  The results of this thesis is that our SDR-based radiometer performs as well as a traditional radiometer and offers additional features that are not found in a traditional radiometer.
